---
title: "Large Seedling Linear Regression"
author: "Kathleen Stutzman"
date: "2024-01-10"
output: html_document
---

Libraries:
```{r Libraries, message=FALSE, warning=FALSE}
# libraries -------------------
library(dplyr)
library(tidyverse)
library(ggplot2)
library(GGally)
library(glmmTMB)
library(TMB)
library(emmeans)
library(DHARMa)
library(lmtest)
library(ggeffects)
library(performance)
```

Set working directory
```{r, setup}
knitr::opts_knit$set(root.dir = '/Users/user/Desktop/Data/Regen_RProj/')
```

Functions:
```{r}
# functions -------------------

```

Source large seedling data:
```{r, message=FALSE, warning=FALSE}
source("Scripts/LS_Import.R")
```

Pivot wider to create dataframe where each row is for one plot and has total details for each species group
```{r}
ls_merge2 <- ls_merge %>% 
  select(Plot_No, Region, Treat_Type, Site, Species_Groups, Total) #this is dropping browse and stump sprout data

ls_merge2 <- ls_merge2 %>% 
  pivot_wider(names_from = Species_Groups, values_from = Total)
```


Import time since data and add it to the large seedling dataset
```{r, message=FALSE}
time_since <- read_csv("CleanData/Treat_Year_Data.csv")

ls_merge3 <- merge(ls_merge2, time_since, by = 'Site')
#log transform time from treatment data
ls_merge3$l.TFT <- log(ls_merge3$Time_from_Treat)
```

Run the 'Add_BA' script and merge with dataset:
```{r, message=FALSE, warning=FALSE}
source("Scripts/Add_BA.R")

# merge with ls dataset -------------------
ls_merge4 <- merge(ls_merge3, prism_BA, by = "Plot_No")
```

Run 'Ground_Data.R' script and add it to large seedling dataset:
```{r}
source("Scripts/Ground_Data.R")

# merge with ls dataset -------------------
ls_merge5 <- merge(ls_merge4, ground3, by = "Plot_No")
```

Source and add veg data
```{r}
source("Scripts/Veg_Data.R")

# merge with ls dataset 
ls.all <- merge(ls_merge5, veg3, by = "Plot_No")

rm(ls_merge5,
   ls_merge2,
   ls_merge3,
   ls_merge4)
```

The large seedling count data is taken in 10m^2^ plots; basal area is measured in hectares; veg and soil data is taken in 1m^2^ plots. 

Large seedling data will be converted into 1m^2^ plots in order to compare across and reduce the amount of scales of data collection to two: 1m^2^ plots and per hectare observations.

Create log transformed categories for newly added variables, then select for just the desired variables:
```{r, eval=FALSE}
ls.all$PIRI.1m <- ls.all$PIRI/10
ls.all$SO.1m <- ls.all$Shrub_Oak/10
ls.all$Other.1m <- ls.all$Other/10

ls.all$l.PIRI1 <- log(ls.all$PIRI.1m + 1)
ls.all$l.SO1 <- log(ls.all$SO.1m + 1)
ls.all$l.other1 <- log(ls.all$Other.1m + 1)

ls.all2 <- ls.all %>% 
  select(Treat_Type, Region, Site, Plot_No, PIRI, PIRI.1m, l.PIRI1, Shrub_Oak, SO.1m, l.SO1, Other, Other.1m, l.other1, Time_from_Treat, l.TFT, BA_HA, l.BA_HA, PIRI.BA_HA, l.BA_piri, Mineral_Soil, l.Mineral, Litter_Duff, avgLD, avgLD_l, Veg_Total, l.Veg_Total) %>% 
  arrange(Treat_Type)
```

Select just for numerical vs log and then look at paired plots:
```{r, eval=FALSE}
#not transformed
ls.num <- ls.all2 %>% 
  select(PIRI, Shrub_Oak, Other, Time_from_Treat, BA_HA, PIRI.BA_HA, Mineral_Soil, avgLD, Veg_Total, Treat_Type)

ggpairs(ls.num)
ggpairs(ls.num, aes(color = Treat_Type))


#log transformed
ls.numl <- ls.all2 %>% 
  select(l.PIRI, l.SO, l.other, l.TFT, l.BA_HA, l.BA_piri, l.Mineral, avgLD_l, l.Veg_Total, Treat_Type)

ggpairs(ls.numl)
ggpairs(ls.numl, aes(color = Treat_Type))

rm(ls.num,
   ls.numl)

rm(ls.all2)
```
Can see the correlation coefficients for linear (Pearsons) relationships. None of them appear very strong, except for ones that are analogs (avg LD vs mineral soil exposure; ba/ha vs piri ba/ha)

Log transformed average litter depth and basal area per hectare have a weak relationship (corr 0.33), which does make sense. 

# The above script a data naming conventions are the same as in the LS_GLMM script




## PIRI Modeling
Best Model AIC: 282.4

I tried with variables at 1m scale - didn't work well.
I also did model elimination not using treatment type - that also did not work well


Revised data set with LS observations at 10m^2^ scale; this means no non-interger values for the poisson distro
```{r}
ls.all3 <- ls.all

ls.all3$l.PIRI <- log(ls.all3$PIRI + 1)
ls.all3$l.SO <- log(ls.all3$Shrub_Oak + 1)
ls.all3$l.other <- log(ls.all3$Other + 1)

ls.all3 <- ls.all3 %>% 
  select(Treat_Type, Region, Site, Plot_No, PIRI, l.PIRI, Shrub_Oak, l.SO, Other, l.other, Time_from_Treat, l.TFT, BA_HA, l.BA_HA, PIRI.BA_HA, l.BA_piri, Mineral_Soil, l.Mineral, Litter_Duff, avgLD, avgLD_l, Veg_Total, l.Veg_Total) %>% 
  arrange(Treat_Type)
```





### testing zero inflation
```{r}
tapply(ls.all3$PIRI, ls.all3$Region, summary)
# no PIRI in ME
tapply(ls.all3$PIRI, ls.all3$Treat_Type, summary) 
```



I started modeling with poisson distro and excluding treatment type. Models without treatment type failed in model fit. Going to run variable elimination again, starting with models with treatment type
```{r}
ls.m13 <- glmmTMB(PIRI ~ Treat_Type + l.SO + l.other + l.BA_HA + l.BA_piri + l.Mineral + avgLD_l + l.Veg_Total + offset(l.TFT) + (1|Site/Plot_No),
                 data = ls.all3,
                 family = nbinom1)
AIC(ls.m13) #288.9

ls.m14 <- glmmTMB(PIRI ~ Treat_Type + l.SO + l.other + l.BA_HA + l.Mineral + avgLD_l + l.Veg_Total + offset(l.TFT) + (1|Site/Plot_No),
                 data = ls.all3,
                 family = nbinom1)
AIC(ls.m14) #287.8

ls.m15 <- glmmTMB(PIRI ~ Treat_Type + l.SO + l.other + l.Mineral + avgLD_l + l.Veg_Total + offset(l.TFT) + (1|Site/Plot_No),
                 data = ls.all3,
                 family = nbinom1)
AIC(ls.m15) #285.9

ls.m16 <- glmmTMB(PIRI ~ Treat_Type + l.SO + l.other + avgLD_l + l.Veg_Total + offset(l.TFT) + (1|Site/Plot_No),
                 data = ls.all3,
                 family = nbinom1)
AIC(ls.m16) #284.4

ls.m17 <- glmmTMB(PIRI ~ Treat_Type + l.SO + avgLD_l + l.Veg_Total + offset(l.TFT) + (1|Site/Plot_No),
                 data = ls.all3,
                 family = nbinom1)
AIC(ls.m17) #283.9

lrtest(ls.m13, ls.m14, ls.m15, ls.m16, ls.m17) #non significant


ls.m18 <- glmmTMB(PIRI ~ Treat_Type + avgLD_l + l.Veg_Total + offset(l.TFT) + (1|Site/Plot_No),
                 data = ls.all3,
                 family = nbinom1)
AIC(ls.m18) #284.4

lrtest(ls.m17, ls.m18) #p = 0.12

ls.m19 <- glmmTMB(PIRI ~ Treat_Type + avgLD_l + offset(l.TFT) + (1|Site/Plot_No),
                 data = ls.all3,
                 family = nbinom1)
summary(ls.m19) #282.4

lrtest(ls.m18, ls.m19) #0.85

# this works, but ls.m19 has lower AIC
ls.m20 <- glmmTMB(PIRI ~ Treat_Type + avgLD_l + offset(l.TFT) + (1|Site/Plot_No),
                 data = ls.all3,
                 ziformula = ~Region,
                 family = nbinom1)
AIC(ls.m20) #286.6


rm(ls.m13, ls.m14, ls.m15, ls.m16, ls.m17, ls.m18)
```



Ok, test model fit with treat type (same as model 11)
```{r}
#Best model
ls.m19 <- glmmTMB(PIRI ~ Treat_Type + avgLD_l + offset(l.TFT) + (1|Site/Plot_No),
                 data = ls.all3,
                 family = nbinom1)
summary(ls.m19) #282.4

ls.m19_sr <- simulateResiduals(ls.m19, n = 1000, plot = TRUE) #passes
testResiduals(ls.m19_sr) #pases
testZeroInflation(ls.m19_sr) #passes
testQuantiles(ls.m19_sr) #passes

#This is a better fit (using nbinom instead of poisson). Convergence issues for nbinom2. Would different distribution change the variables included? - No, I went back and ran model elimination with nbinom1 distro instead of poisson. same variables are significant

ls.m20_sr <- simulateResiduals(ls.m20, n = 1000, plot = TRUE) #passes
testResiduals(ls.m20_sr) #passes
testZeroInflation(ls.m20_sr) #passes
testQuantiles(ls.m20_sr) #passes


```




Trying pairwise comparison of model fit:
```{r}
emmeans(ls.m19, specs = pairwise ~ Treat_Type, adjust = 'Tukey', type = 'response')
```

Control vs Harvest is the only significant different (p = 0.0271)



Would things fit better with zi ~Region? Answer: NO, good model fit but higher AIC
```{r, include=FALSE, eval=FALSE}
ls.m13a <- glmmTMB(PIRI ~ Treat_Type + l.SO + l.other + l.BA_HA + l.BA_piri + l.Mineral + avgLD_l + l.Veg_Total + offset(l.TFT) + (1|Site/Plot_No),
                 data = ls.all3,
                 ziformula = ~Region,
                 family = nbinom1)
AIC(ls.m13a) #doesn't converge

#test piri ba
ls.m14a <- glmmTMB(PIRI ~ Treat_Type + l.SO + l.other + l.BA_HA + l.Mineral + avgLD_l + l.Veg_Total + offset(l.TFT) + (1|Site/Plot_No),
                 data = ls.all3,
                 ziformula = ~Region,
                 family = nbinom1)
AIC(ls.m14a) #291.7

#test mineral
ls.m15a <- glmmTMB(PIRI ~ Treat_Type + l.SO + l.other + l.BA_HA + avgLD_l + l.Veg_Total + offset(l.TFT) + (1|Site/Plot_No),
                 data = ls.all3,
                 ziformula = ~Region,
                 family = nbinom1)
AIC(ls.m15a) #290.3

#test so
ls.m16a <- glmmTMB(PIRI ~ Treat_Type + l.other + l.BA_HA + avgLD_l + l.Veg_Total + offset(l.TFT) + (1|Site/Plot_No),
                 data = ls.all3,
                 ziformula = ~Region,
                 family = nbinom1)
AIC(ls.m16a) #290.7

#test other
ls.m17a <- glmmTMB(PIRI ~ Treat_Type + l.BA_HA + avgLD_l + l.Veg_Total + offset(l.TFT) + (1|Site/Plot_No),
                 data = ls.all3,
                 ziformula = ~Region,
                 family = nbinom1)
AIC(ls.m17a) #289.7

lrtest(ls.m13a, ls.m14a, ls.m15a, ls.m16a, ls.m17a) #non significant

#test BA
ls.m18a <- glmmTMB(PIRI ~ Treat_Type + avgLD_l + l.Veg_Total + offset(l.TFT) + (1|Site/Plot_No),
                 data = ls.all3,
                 ziformula = ~Region,
                 family = nbinom1)
AIC(ls.m18a) #288.6

lrtest(ls.m17a, ls.m18a) #p = 0.12

#test veg - best model
ls.m19a <- glmmTMB(PIRI ~ Treat_Type + avgLD_l + offset(l.TFT) + (1|Site/Plot_No),
                 data = ls.all3,
                 ziformula = ~Region,
                 family = nbinom1)
AIC(ls.m19a) #286.6

# test ld
ls.m20a <- glmmTMB(PIRI ~ Treat_Type + offset(l.TFT) + (1|Site/Plot_No),
                 data = ls.all3,
                 ziformula = ~Region,
                 family = nbinom1)
AIC(ls.m20a) #294.9
lrtest(ls.m19a, ls.m20a) # p = 0.002

rm(ls.m13a, ls.m14a, ls.m15a, ls.m16a, ls.m17a, ls.m18a, ls.m20a)



ls.m19a_sr <- simulateResiduals(ls.m19a, n = 1000, plot = T) #passes, but AIC is higher than without zi
testResiduals(ls.m19a_sr) #passes

rm(ls.m19a_sr, ls.m19a)
```








#### GG Plot for ls piri
```{r, fig.align='center'}
# PIRI LS -------------------
ls.p1 <- ggpredict(ls.m19, terms = c("avgLD_l", "Treat_Type"))


ggplot(ls.p1, aes(x, predicted, color = group))+
  geom_line(linewidth = 1.25, show.legend = FALSE) +
  scale_color_manual(values = c("#D8B70A",  "#02401B", "#A2A475", "#81A88D", "#972D15"))+
  theme_classic()+
  theme(panel.background = element_blank()) +
  theme(axis.text = element_text(size = 16), axis.title = element_text(size = 20))+
  labs(x = "Average leaf litter depth \n (log transformed)",
       y = "Pitch pine stems \n (adjusted for time)")

ggsave(filename = "LS_PIRI_avgld.tiff", path = "Plots/PIRI_GLMM", width = 7, height = 5, device = "tiff", dpi = 700)
```









# Shrub oak models
Best Model AIC: 3267
Issue with underdispersion

#### Are SO LS present in all regions and treatment types?
Yes   
```{r}
tapply(ls.all3$Shrub_Oak, ls.all3$Region, summary)
tapply(ls.all3$Shrub_Oak, ls.all3$Treat_Type, summary)
```

Tried a lot of distributions. Models are generally underdispersed. Have used poisson, nbinom, genpois, & compois distros. Using zi ~1 helps. 
```{r, eval=FALSE}
# so.ls1 <- glmmTMB(Shrub_Oak ~ Treat_Type + l.PIRI + l.other + l.BA_HA + l.BA_piri + l.Mineral + avgLD_l + l.Veg_Total + offset(l.TFT) + (1|Site/Plot_No),
#                  data = ls.all3,
#                  ziformula = ~1,
#                  family = genpois)
#AIC(so.ls1) # fails with genpois

# test piri ba
so.ls2 <- glmmTMB(Shrub_Oak ~ Treat_Type + l.PIRI + l.other + l.BA_HA + l.Mineral + avgLD_l + l.Veg_Total + offset(l.TFT) + (1|Site/Plot_No),
                 data = ls.all3,
                 ziformula = ~1,
                 family = genpois)
AIC(so.ls2) #3266.7

#lrtest(so.ls1, so.ls2) #p = 0.3

# test total BA
so.ls3 <- glmmTMB(Shrub_Oak ~ Treat_Type + l.PIRI + l.other + l.Mineral + avgLD_l + l.Veg_Total + offset(l.TFT) + (1|Site/Plot_No),
                 data = ls.all3,
                 ziformula = ~1,
                 family = genpois)
AIC(so.ls3) #3268

lrtest(so.ls2, so.ls3) # p = 0.06

# test mineral
so.ls4 <- glmmTMB(Shrub_Oak ~ Treat_Type + l.PIRI + l.other + avgLD_l + l.Veg_Total + offset(l.TFT) + (1|Site/Plot_No),
                 data = ls.all3,
                 ziformula = ~1,
                 family = genpois)
AIC(so.ls4) #3268.6


lrtest(so.ls3, so.ls4) # p = 0.1

# test avg ld
so.ls5 <- glmmTMB(Shrub_Oak ~ Treat_Type + l.PIRI + l.other + l.Veg_Total + offset(l.TFT) + (1|Site/Plot_No),
                 data = ls.all3,
                 ziformula = ~1,
                 family = genpois)
AIC(so.ls5) #3267.3

lrtest(so.ls5, so.ls4) # p = 0.4

# test piri*
so.ls6 <- glmmTMB(Shrub_Oak ~ Treat_Type + l.other + l.Veg_Total + offset(l.TFT) + (1|Site/Plot_No),
                 data = ls.all3,
                 ziformula = ~1,
                 family = genpois)
AIC(so.ls6) #fails

lrtest(so.ls6, so.ls5) # p = 0.008**

# test other*
so.ls7 <- glmmTMB(Shrub_Oak ~ Treat_Type + l.Veg_Total + l.PIRI + offset(l.TFT) + (1|Site/Plot_No),
                 data = ls.all3,
                 ziformula = ~1,
                 family = genpois)
AIC(so.ls7) #fails

lrtest(so.ls5, so.ls7) # p = 0.003

# test veg*
so.ls8 <- glmmTMB(Shrub_Oak ~ Treat_Type +l.other + l.PIRI + offset(l.TFT) + (1|Site/Plot_No),
                 data = ls.all3,
                 ziformula = ~1,
                 family = genpois)
AIC(so.ls8) # 3271.1

# test piri ba, model failed at start
so.ls9 <- glmmTMB(Shrub_Oak ~ Treat_Type +l.other + l.PIRI + l.Veg_Total + l.BA_piri + offset(l.TFT) + (1|Site/Plot_No),
                 data = ls.all3,
                 ziformula = ~1,
                 family = genpois)
AIC(so.ls9) #3268.4

lrtest(so.ls9, so.ls5) # p = 0.4
```


Best Model
```{r}
so.ls5 <- glmmTMB(Shrub_Oak ~ Treat_Type + l.PIRI + l.other + l.Veg_Total + offset(l.TFT) + (1|Site/Plot_No),
                 data = ls.all3,
                 ziformula = ~1,
                 family = genpois)
summary(so.ls5)
AIC(so.ls5)

so.ls5_sr <- simulateResiduals(so.ls5, n = 1000, plot = T)
testResiduals(so.ls5_sr)
testDispersion(so.ls5_sr, alternative = "less")

check_collinearity(so.ls5) #low

```



### Pairwise with treatment type
```{r, eval=FALSE}
emmeans(so.ls5, specs = pairwise ~ Treat_Type, adjust = 'Tukey', type = 'response')
```
# FallRx, Harvest, MowRx, & SpringRx all significantly different from Control. Only Harvest & MowRx significantly different from each outher







### testing other distros
```{r, eval=FALSE}
so.ls6 <- glmmTMB(Shrub_Oak ~ Treat_Type + l.other + l.BA_HA + l.Mineral + l.Veg_Total + offset(l.TFT) + (1|Site/Plot_No),
                 data = ls.all3,
                 family = poisson)
AIC(so.ls6)

so.ls6a <- glmmTMB(Shrub_Oak ~ Treat_Type + l.other + l.BA_HA + l.Mineral + l.Veg_Total + offset(l.TFT) + (1|Site/Plot_No),
                 data = ls.all3,
                 family = genpois)
AIC(so.ls6a)

# compois distro fails, won't produce DHARMa results

so.ls6b <- glmmTMB(Shrub_Oak ~ Treat_Type + l.other + l.BA_HA + l.Mineral + l.Veg_Total + offset(l.TFT) + (1|Site/Plot_No),
                 data = ls.all3,
                 family = tweedie)
AIC(so.ls6b)

so.ls6b_sr <- simulateResiduals(so.ls6b, n = 1000, plot = TRUE)
testDispersion(so.ls6b_sr) #fails

so.ls6c <- glmmTMB(Shrub_Oak ~ Treat_Type + l.other + l.BA_HA + l.Mineral + l.Veg_Total + offset(l.TFT) + (1|Site/Plot_No),
                 data = ls.all3,
                 family = t_family)
AIC(so.ls6c)

so.ls6c_sr <- simulateResiduals(so.ls6c, n = 1000, plot = TRUE)
testResiduals(so.ls6c_sr)
testDispersion(so.ls6c_sr) #passes
testQuantiles(so.ls6c_sr)


so.ls6d <- glmmTMB(Shrub_Oak ~ Treat_Type + l.other + l.BA_HA + l.Mineral + l.Veg_Total + offset(l.TFT) + (1|Site/Plot_No),
                 data = ls.all3,
                 family = gaussian)
AIC(so.ls6d)

so.ls6d_sr <- simulateResiduals(so.ls6d, n = 1000, plot = TRUE) #fails
plot(so.ls6d_sr)
testResiduals(so.ls6d_sr)
```


```{r}
hist(ls.all3$Shrub_Oak)

ggplot(ls.all3, aes(x=Shrub_Oak))+
  geom_histogram(binwidth = 2)+
  facet_grid(rows = vars(Treat_Type))
```



### Graph model

```{r, fig.align='center'}
# SO LS -------------------
ls.s1 <- ggpredict(so.ls5, terms = c("l.PIRI", "Treat_Type"))


ggplot(ls.s1, aes(x, predicted, color = group))+
  geom_line(linewidth = 1.25) +
  scale_color_manual(values = c("#D8B70A",  "#02401B", "#A2A475", "#81A88D", "#972D15"))+
  theme_classic()+
  theme(panel.background = element_blank()) +
  theme(axis.text = element_text(size = 16), axis.title = element_text(size = 20))+
  labs(x = "Count of pitch pine seedlings (log)",
       y = "Shrub oak stems \n(adjusted for time)")

#ggsave(filename = "LS_PIRI_avgld.tiff", path = "Plots/PIRI_GLMM", width = 7, height = 5, device = "tiff", dpi = 700)


ls.s2 <- ggpredict(so.ls5, terms = c("l.other", "Treat_Type"))


ggplot(ls.s2, aes(x, predicted, color = group))+
  geom_line(linewidth = 1.25, show.legend = FALSE) +
  scale_color_manual(values = c("#D8B70A",  "#02401B", "#A2A475", "#81A88D", "#972D15"))+
  theme_classic()+
  theme(panel.background = element_blank()) +
  theme(axis.text = element_text(size = 16), axis.title = element_text(size = 20))+
  labs(x = "Count of other seedlings (log)",
       y = "Shrub oak stems \n(adjusted for time)")

# -------------------
ls.s3 <- ggpredict(so.ls5, terms = c("l.Veg_Total", "Treat_Type"))


ggplot(ls.s3, aes(x, predicted, color = group))+
  geom_line(linewidth = 1.25, show.legend = FALSE) +
  scale_color_manual(values = c("#D8B70A",  "#02401B", "#A2A475", "#81A88D", "#972D15"))+
  theme_classic()+
  theme(panel.background = element_blank()) +
  theme(axis.text = element_text(size = 16), axis.title = element_text(size = 20))+
  labs(x = "Total understory vegetative cover (log)",
       y = "Shrub oak stems \n(adjusted for time)")

# low collinearity but still gives me a little pause..

```
## Interesting - it is always Mowing -> SpringRx -> Fall Rx -> Harvest -> Control




I've tried poisson, nbinom1, nbinom1 with ZI, nbinom2, genpois, compois, tweedie, gaussian and none of these distros work. They are all underdispersed


### I've deleted more models than you've imagined.  
Genpois with zi ~1 seems like the closest I'm going to get















