---
title: "Large seedling models for Tony & Kevin"
author: "Kathleen Stutzman"
date: "2024-01-25"
output: html_document
---

Hidden below is code for importing the data set and log transforming variables. The best model (lowest AIC & using lrtest) is presented below, model fit checked, and pairwise comparison for Treatment Type.  

I did create and test models without treatment type, but they always have higher AIC and worse model fit that those with (probably due to the control category...)  

##### Variables investigated:
- **Treat_Type** - treatment type
- **PIRI** - pitch pine seedling count (10m^2^)
- **SO** - shrub oak seedling count (10m^2^)
- **other** - other seedling count (10m^2^)
- **BA_HA** - basal area per hectare
- **BA_piri** - pitch pine basal area per hectare
- **Mineral** - exposed mineral soil (1m^2^)
- **avgLD** - average leaf litter depth (1m^2^)
- **Veg_Total** - total understory plant cover (1m^2^)
- **TFT** - time between treatment and sampling (used as an offset)
- **Site/Plot_No** - random effect of plots within sites

*Variables are log transformed (l. or _l)*

## Pitch Pine Regeneration

```{r Libraries, message=FALSE, warning=FALSE, include=FALSE}
# libraries -------------------
library(dplyr)
library(tidyverse)
library(ggplot2)
library(GGally)
library(glmmTMB)
library(TMB)
library(emmeans)
library(DHARMa)
library(lmtest)
```

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = '/Users/user/Desktop/Data/Regen_RProj/')
```

```{r, message=FALSE, warning=FALSE, include=FALSE}
source("Scripts/LS_Import.R")

ls_merge2 <- ls_merge %>% 
  select(Plot_No, Region, Treat_Type, Site, Species_Groups, Total) #this is dropping browse and stump sprout data

ls_merge2 <- ls_merge2 %>% 
  pivot_wider(names_from = Species_Groups, values_from = Total)

time_since <- read_csv("CleanData/Treat_Year_Data.csv")

ls_merge3 <- merge(ls_merge2, time_since, by = 'Site')
#log transform time from treatment data
ls_merge3$l.TFT <- log(ls_merge3$Time_from_Treat)

source("Scripts/Add_BA.R")

# merge with ls dataset -------------------
ls_merge4 <- merge(ls_merge3, prism_BA, by = "Plot_No")

source("Scripts/Ground_Data.R")

# merge with ls dataset -------------------
ls_merge5 <- merge(ls_merge4, ground3, by = "Plot_No")

source("Scripts/Veg_Data.R")

# merge with ls dataset 
ls.all <- merge(ls_merge5, veg3, by = "Plot_No")

rm(ls_merge5,
   ls_merge2,
   ls_merge3,
   ls_merge4)

ls.all3 <- ls.all

ls.all3$l.PIRI <- log(ls.all3$PIRI + 1)
ls.all3$l.SO <- log(ls.all3$Shrub_Oak + 1)
ls.all3$l.other <- log(ls.all3$Other + 1)

ls.all3 <- ls.all3 %>% 
  select(Treat_Type, Region, Site, Plot_No, PIRI, l.PIRI, Shrub_Oak, l.SO, Other, l.other, Time_from_Treat, l.TFT, BA_HA, l.BA_HA, PIRI.BA_HA, l.BA_piri, Mineral_Soil, l.Mineral, Litter_Duff, avgLD, avgLD_l, Veg_Total, l.Veg_Total) %>% 
  arrange(Treat_Type)
```

### Best large seedling pitch pine model
```{r}
ls.m19 <- glmmTMB(PIRI ~ Treat_Type + avgLD_l + offset(l.TFT) + (1|Site/Plot_No),
                 data = ls.all3,
                 family = poisson)
summary(ls.m19)
```

### Check model fit
```{r}
ls.m19_sr <- simulateResiduals(ls.m19, n = 1000, plot = TRUE) #passes KS, quantile deviations fails, but it still could be an accepted model
testResiduals(ls.m19_sr) #passes
testZeroInflation(ls.m19_sr) #passes
testQuantiles(ls.m19_sr) #fails
```
Issues with quantiles, but still can be a good/useful model despite these issues


### Pairwise with treatment type
##### These are really rates due to the offset for time
```{r}
emmeans(ls.m19, specs = pairwise ~ Treat_Type, adjust = 'Tukey', type = 'response')
```
##### Only harvest is significantly different than contol


### Graph pitch pine models
```{r, warning=FALSE, message=FALSE, fig.align='center'}
library(sjPlot)
library(sjlabelled)
library(sjmisc)

set_theme(base = theme_classic(),
          theme.font = 'serif',
          axis.title.size = 1.5,
          axis.textsize.x = 1.5,
          axis.textsize.y = 1.5,
          title.size = 2.5,
          title.align = "center",
          legend.pos = "right",
          legend.size = 1.5,
          legend.title.size = 1.5,
          #legend.bordercol = "black",
          legend.item.size = .75)


# LS PIRI plot 1 ------------------- LD vs. TT
plot_model(ls.m19, 
                       type = 'pred', 
                       terms = c('avgLD_l', 'Treat_Type'),
                       axis.title = c("Average Leaf Litter Depth (log transformed)", "Total Count of Pitch Pine"),
                       title = "Predicted Counts of Pitch Pine \n Seedlings >/=50cm & <2.5cm DBH",
           legend.title = "Treatment Type",
           line.size = 1,
           value.offset = 'Treat_Type',
           ci.lvl = NA,
           colors = c("#D8B70A", "#02401B", "#A2A475", "#81A88D", "#972D15")) 




# LS PIRI plot 2 ------------------- Time from Treatment vs TT
plot_model(ls.m19, 
                       type = 'pred', 
                       terms = c('l.TFT', 'Treat_Type'),
                       axis.title = c("Time from Treatment (log transformed)", "Total Count of Pitch Pine"),
                       title = "Predicted Counts of Pitch Pine \n Seedlings >/=50cm & <2.5cm DBH",
           legend.title = "Treatment Type",
           line.size = 1,
           value.offset = 'Treat_Type',
           ci.lvl = NA,
           colors = c("#D8B70A", "#02401B", "#A2A475", "#81A88D", "#972D15")) 

```


## Shrub Oak Regeneration
This includes QUIL (scrub oak) & QUPR (dwarf chinkapin oak)

### Large seedling shrub oak model
  - Poisson, genpois, & compois models all fail and are underdispersed. Gaussian dispersion also fails. The student t distro seems to work ...
  - Model versions suggest that total veg cover is important. However, shrub oak can often make up a significant portion of understory veg cover. So, of course shrub oak abundance increases as total veg cover increases. Therefore, it is dropped from the model for issues of collineraity. 
```{r}
ls.all3$BA_10m <- (ls.all3$BA_HA/1000)
ls.all3$l.BA_10m <- log(ls.all3$BA_10m + 1)

so.ls6 <- glmmTMB(Shrub_Oak ~ Treat_Type + l.other + l.BA_10m + l.Mineral + offset(l.TFT) + (1|Site/Plot_No),
                 data = ls.all3,
                 family = t_family)
summary(so.ls6)
```

### Check model fit
```{r}
so.ls6_sr <- simulateResiduals(so.ls6, n = 1000, plot = TRUE)
testResiduals(so.ls6_sr)
```
A little concerned about the DHARMa nonparametric dispersion test graph

### Pairwise with treatment type
```{r}
emmeans(so.ls6, specs = pairwise ~ Treat_Type, adjust = 'Tukey', type = 'response')
```
#### Seems like MowRx is signficantly different from all other categories


### Graph shrub oak model
```{r}
# LS SO plot 1 ------------------- other vs. TT
plot_model(so.ls6, 
                       type = 'pred', 
                       terms = c('l.other', 'Treat_Type'),
                       axis.title = c("Other small seedling counts (log transformed)", "Total Count of Shrub Oak"),
                       title = "Predicted Counts of Shrub Oak \n Seedlings >/=50cm & <2.5cm DBH",
           legend.title = "Treatment Type",
           line.size = 1,
           value.offset = 'Treat_Type',
           ci.lvl = NA,
           colors = c("#D8B70A", "#02401B", "#A2A475", "#81A88D", "#972D15")) 

# LS SO plot 2 ------------------- BA_10m vs. TT
plot_model(so.ls6, 
                       type = 'pred', 
                       terms = c('l.other', 'Treat_Type'),
                       axis.title = c("Total Basal Area per 10 square meters (log transformed)", "Total Count of Shrub Oak"),
                       title = "Predicted Counts of Shrub Oak \n Seedlings >/=50cm & <2.5cm DBH",
           legend.title = "Treatment Type",
           line.size = 1,
           value.offset = 'Treat_Type',
           ci.lvl = NA,
           colors = c("#D8B70A", "#02401B", "#A2A475", "#81A88D", "#972D15")) 

# LS SO plot 3 ------------------- l.mineral vs. TT
plot_model(so.ls6, 
                       type = 'pred', 
                       terms = c('l.other', 'Treat_Type'),
                       axis.title = c("Amount of exposed mineral soil (log transformed)", "Total Count of Shrub Oak"),
                       title = "Predicted Counts of Shrub Oak \n Seedlings >/=50cm & <2.5cm DBH",
           legend.title = "Treatment Type",
           line.size = 1,
           value.offset = 'Treat_Type',
           ci.lvl = NA,
           colors = c("#D8B70A", "#02401B", "#A2A475", "#81A88D", "#972D15")) 
```











