---
title: "Sapling PIRI GLMM"
author: "Kathleen Stutzman"
date: "2024-01-04"
output: html_document
---

## Introduction:
This script, originally written as an R script, will be much more readable as an R Markdown. Therefore, I am translating it into a markdown file.

# Pitch Pine Sapling GLMM code

Libraries:
```{r Libraries, message=FALSE, warning=FALSE}
# libraries -------------------
library(dplyr)
library(tidyverse)
library(ggplot2)
library(GGally)
library(glmmTMB)
library(TMB)
library(emmeans)
library(DHARMa)
```

Set working directory
```{r, setup}
knitr::opts_knit$set(root.dir = '/Users/user/Desktop/Data/Regen_RProj/')
```

Functions:
```{r}
# functions -------------------

```

Source sapling data:
```{r}
source("Scripts/SA_Import.R")
```

Pivot wider to create dataframe where each row is for one plot and has total details for each species group
```{r}
sa_merge2 <- sapling_merge %>% 
  pivot_wider(names_from = Species_Groups, values_from = Total_Tally)
```

Import time since data and add it to the sapling dataset
```{r}
time_since <- read_csv("CleanData/Treat_Year_Data.csv")

sa_merge3 <- merge(sa_merge2, time_since, by = 'Site')
#log transform time from treatment data
sa_merge3$l.TFT <- log(sa_merge3$Time_from_Treat)
```

Run the 'Add_BA' script and merge with dataset:
```{r}
source("Scripts/Add_BA.R")

# merge with sapling dataset -------------------
sa_merge4 <- merge(sa_merge3, prism_BA, by = 'Plot_No')
```

Run 'Ground_Data.R' script and add it to sapling dataset:
```{r}
source("Scripts/Ground_Data.R")

# merge with sapling dataset -------------------
sa_merge5 <- merge(sa_merge4, ground3, by = 'Plot_No')

```

Source and add veg data:
```{r}
source("Scripts/Veg_Data.R")

# merge with sapling dataset -------------------
sa.all <- merge(sa_merge5, veg3, by = "Plot_No")

rm(sa_merge3,
   sa_merge4,
   sa_merge5)
```

Sapling count data is taken in 25m^2^ plots; basal area is measured in hectares; veg and soil data is taken a 1m^2^ plots.

Sapling data will be convered into 1m^2^ plots in order to compare across and reduce the amount of scales of data collection, reducing it to two: 1m^2^ plots and per hectare observations

```{r}
sa.all$PIRI.1m <- sa.all$PIRI/25
sa.all$SO.1m <- sa.all$Shrub_Oak/25
sa.all$Other.1m <- sa.all$Other/25
```


Create log transformed categories for newly added variables, then select for just the desired variables:
```{r}
sa.all$l.PIRI <- log(sa.all$PIRI.1m + 1)
sa.all$l.SO <- log(sa.all$SO.1m + 1)
sa.all$l.other <- log(sa.all$Other.1m + 1)

sa.all2 <- sa.all %>% 
  select(Treat_Type, Region, Site, Plot_No, PIRI, l.PIRI, Shrub_Oak, l.SO, Other, l.other, Time_from_Treat, l.TFT, BA_HA, l.BA_HA, PIRI.BA_HA, l.BA_piri, Mineral_Soil, l.Mineral, Litter_Duff, avgLD, avgLD_l, Veg_Total, l.Veg_Total) %>% 
  arrange(Treat_Type)
```





Select just for numerical vs log and then look at paired plots:
```{r, eval=FALSE}
#not transformed
sa.num <- sa.all2 %>% 
  select(PIRI, Shrub_Oak, Other, Time_from_Treat, BA_HA, PIRI.BA_HA, Mineral_Soil, avgLD, Veg_Total, Treat_Type)

ggpairs(sa.num)
ggpairs(sa.num, aes(color = Treat_Type))


#log transformed
sa.numl <- sa.all2 %>% 
  select(l.PIRI, l.SO, l.other, l.TFT, l.BA_HA, l.BA_piri, l.Mineral, avgLD_l, l.Veg_Total, Treat_Type)

ggpairs(sa.numl)
ggpairs(sa.numl, aes(color = Treat_Type))

rm(sa.num,
   sa.numl)
```
Can see the correlation coefficients for linear (Pearsons) relationships. None of them appear very strong, except for ones that are analogs (avg LD vs mineral soil exposure; ba/ha vs piri ba/ha)



No significant correlations obvious from plots


# Above code and naming conventions are the same as SA_LR.Rmd script

























Hidden below are models with convergence problems:
```{r, eval=FALSE, include=FALSE}
sa.nb_all <- glmer.nb(Total_Tally~Treat_Type + l.BA_HA + l.BA_piri + l.Mineral + avgLD_l + (1|Site/Plot_No) + offset(l.TFT),
                      data = sa_alldata)
summary(sa.nb_all) #failed to converge

sa.nb_all <- glmmTMB(Total_Tally~Treat_Type + l.BA_HA + l.BA_piri + l.Mineral + avgLD_l + (1|Site/Plot_No) + offset(l.TFT),
                      data = sa_alldata,
                     family = nbinom2)
summary(sa.nb_all) #still fails to converge using glmmTMB

sa.nb2 <- glmer.nb(Total_Tally~Treat_Type + l.Mineral + avgLD_l + (1|Site/Plot_No) + offset(l.TFT),
                      data = sa_alldata)
summary(sa.nb2) #failed to converge


rm(sa.nb2, sa.nb_all)
```

Models that converge:
```{r}
sa.nb2 <- glmmTMB(Total_Tally~Treat_Type + l.Mineral + avgLD_l + (1|Site/Plot_No) + offset(l.TFT),
                      data = sa_alldata,
                  family = nbinom2)
AIC(sa.nb2)  #AIC = 217.6 - this does converge

sa.p3 <- glmer(Total_Tally~Treat_Type + avgLD_l + (1|Site/Plot_No) + offset(l.TFT),
               data = sa_alldata,
               family = poisson)
summary(sa.p3) #first model that didn't fail to converge, AIC is 213.6

######## this is the best model
sa.p4 <- glmer(Total_Tally~Treat_Type + (1|Site/Plot_No) + offset(l.TFT),
               data = sa_alldata,
               family = poisson)
summary(sa.p4) #AIC is 211.7

# would like to try this model again, but with all the data (as model is not using veg/soil data, which cuts sample size in half)
sa.p5 <- glmmTMB(Total_Tally~Treat_Type + (1|Site/Plot_No) + offset(l.TFT),
               data = sa.merge1,
               family = poisson)
AIC(sa.p5) #409.1 - converges using glmmTMB

sa.p6 <- glmer(Total_Tally~Treat_Type + (1|Site/Plot_No),
               data = sa_alldata,
               family = poisson)
summary(sa.p6) #AIC is 211.3

sa.p7 <- glmer(Total_Tally~Treat_Type + l.BA_piri + (1|Site/Plot_No) + offset(l.TFT),
               data = sa_alldata,
               family = poisson)
summary(sa.p7)
```

Test models:
```{r}
library(lmtest)
lrtest(sa.p4, sa.p6) #same p-value as the offset in other tests, wonder what that's about

lrtest(sa.p4, sa.p7) #p = .2, not worth the l.BA_piri

rm(sa.p3, sa.p6, sa.p7)
```

Run DHARMa package to check model fit:
```{r}
sa.p4_sr <- simulateResiduals(fittedModel = sa.p4, n=1000, plot = TRUE)

testDispersion(sa.p4_sr)
testZeroInflation(sa.p4_sr)

#no good, model is zero inflated
```

Data is zero inflated. Need to try zero inflated models:
```{r}
######## this is the best model
sa.zi4 <- glmmTMB(Total_Tally~Treat_Type + (1|Site/Plot_No) + offset(l.TFT),
               data = sa_alldata,
               ziformula=~1,
               family = poisson)
summary(sa.zi4) #AIC is 257.7 - this passes the DHARMa test

sa.zi4_sr <- simulateResiduals(fittedModel = sa.zi4, n = 1000, plot = TRUE)

testDispersion(sa.zi4_sr)# PASSES
testOutliers(sa.zi4_sr)
testZeroInflation(sa.zi4_sr) #looking a lot better

plotResiduals(sa.zi4_sr, sa_alldata$Treat_Type, xlab = "Treatment Type", main=NULL)  

#plotResiduals(sa.zi4_sr, sa_alldata$Site, xlab = "Site", main=NULL)

# would like to try this model again, but with all the data
sa.zi9 <- glmmTMB(Total_Tally~Treat_Type + (1|Site/Plot_No) + offset(l.TFT),
                  data = sa.merge1,
                  ziformula=~1,
                  family = poisson)
summary(sa.zi9) #AIC is 411.1 - lets check with dharma 

sa.zi9_sr <- simulateResiduals(fittedModel = sa.zi9, n = 1000, plot = TRUE) #significant differences

testDispersion(sa.zi9_sr)# does not pass
testOutliers(sa.zi9_sr)
testZeroInflation(sa.zi9_sr) #does not pass
#looks like it is better to use half the data, which is what i'm using for the other two measurements, so at least it is consistent

rm(sa.zi9,
   sa.zi9_sr)
```


Here are zero-inflated models that converge but do not pass DHARMa tests
```{r, eval=FALSE}
sa.zi5 <- glmmTMB(Total_Tally~Treat_Type + l.BA_HA + (1|Site/Plot_No) + offset(l.TFT),
                  data = sa_alldata,
                  ziformula=~1,
                  family = poisson)
summary(sa.zi5) #AIC is 215.4, SR don't agree - not better!


sa.zi6 <- glmmTMB(Total_Tally~Treat_Type + l.BA_piri + (1|Site/Plot_No) + offset(l.TFT),
                  data = sa_alldata,
                  ziformula=~1,
                  family = poisson)
summary(sa.zi6) #AIC is 214.2, SR don't agree - not better!

sa.zi7 <- glmmTMB(Total_Tally~Treat_Type + l.BA_piri + (1|Site/Plot_No),
                  data = sa_alldata,
                  ziformula=~1,
                  family = poisson)
summary(sa.zi7) #AIC is 213.9, SR don't agree - not better!

sa.zi8 <- glmmTMB(Total_Tally~Treat_Type + (1|Site/Plot_No),
                  data = sa_alldata,
                  ziformula=~1,
                  family = poisson)
summary(sa.zi8) #doesn't pass the DHARMa test

rm(sa.zi5,
   sa.zi6,
   sa.zi7,
   sa.zi8)
```

Use emmeans for pairwise comparison of Treatment Types:
```{r}
emmeans(sa.zi4, specs = pairwise ~ Treat_Type, adjust = 'Tukey', type = 'response')
#none significant
```

Below is script to create plots. Do not run unless desired. The added packages cause issues for running general GLMM scripts (issues with tidyverse) and therefore are only worth activating if you want to produce graphs or have already loaded all desired analysis
```{r, eval=FALSE}
library(sjPlot)
library(sjlabelled)
library(sjmisc)

set_theme(base = theme_classic(),
          theme.font = 'serif',
          axis.title.size = 1.5,
          axis.textsize.x = 1.5,
          axis.textsize.y = 1.5,
          title.size = 2.5,
          title.align = "center",
          legend.pos = "right",
          legend.size = 1.5,
          legend.title.size = 1.5,
          #legend.bordercol = "black",
          legend.item.size = .75)

plot_model(sa.zi4, 
           type = "pred",
           terms = "Treat_Type")

plot_model(sa.zi4, 
           type = "pred",
           terms = c("l.TFT", "Treat_Type"),
           axis.title = c("Time from Last Treatment (log transformed)", "Total Count of Pitch Pine"),
           title = "Predicted Counts of Pitch Pine Saplings \n >/= 2.5 cm & < 10 cm DBH",
           legend.title = "Treatment Type",
           line.size = 1,
           show.zeroinf = T,
           ci.lvl = NA,
           colors = c("#D8B70A", "#02401B", "#A2A475", "#81A88D", "#972D15"))
```


# Other hidden analysis: early models
```{r, include=FALSE, eval=FALSE}
##################### this is my original analysis of all plot before I added more data / variables to the model -------------------
sa.nb_null <- glmer.nb(Total_Tally~Treat_Type + (1|Site/Plot_No),
                       data = sapling_PIRI)
summary(sa.nb_null) #AIC is 434

sa.zi_null <- glmmTMB(Total_Tally~Treat_Type + (1|Site/Plot_No),
                       ziformula=~1,
                       family = nbinom2,
                       data=sapling_PIRI)
summary(sa.zi_null) #AIC is higher - need to learn more about how to fit/check the fit of models


# using DHARMa to check the model fit -------------------

sa_sim.resid <- simulateResiduals(sa.nb_null, n = 1000, plot = TRUE) #model differs significatly

testDispersion(sa_sim.resid)
testOutliers(sa_sim.resid)

plotResiduals(sa_sim.resid, sapling_PIRI$Treat_Type, xlab = "Treatment_Type", main=NULL)

plotResiduals(sa_sim.resid, sapling_PIRI$Site, xlab = "Site", main=NULL)

outliers(sa_sim.resid) #integer(0) is the given response???

testZeroInflation(sa_sim.resid) #this data appears to be zero inflated, so I have to figure out more about writing zero inflated code and then recheck

sa.zi_region <- glmmTMB(Total_Tally~Treat_Type + (1|Site/Plot_No),
                      zi=~Region,
                      family = nbinom2,
                      data=sapling_PIRI)
summary(sa.zi_region) #AIC is 467.7, more than null model, but null model is a poor fit
#nbinom1 as family has model convergence problems

sa.zi_sim.resid <- simulateResiduals(sa.zi_region, n=1000, plot = TRUE)

testZeroInflation(sa.zi_sim.resid) #holy shit, this looks good


# plot residuals against predictors/random variables/zi assignment
plotResiduals(sa.zi_sim.resid, sapling_PIRI$Treat_Type, xlab = "Treat_Type", main = NULL)

plotResiduals(sa.zi_sim.resid, sapling_PIRI$Site, xlab = "Site", main = NULL)

plotResiduals(sa.zi_sim.resid, sapling_PIRI$Region, xlab = "Region", main = NULL)

# post-hoc pairwise comparison -------------------
emmeans(sa.zi_region, pairwise~Treat_Type, adjust = "Tukey", type = "response") #not significant
```




# Hidden analysis: filter just for PIRI
```{r, include=FALSE, eval=FALSE}
# filter for just piri -------------------
sapling_PIRI <- sapling_merge %>% 
  filter(Species_Groups == "PIRI") %>% 
  arrange(Treat_Type)
# PIRI saplings are only in 50 of the 1001 plots and mostly in control sites....


# now to make the explanatory variables factors
sapling_PIRI$Treat_Type = factor(sapling_PIRI$Treat_Type,
                              levels = unique(sapling_PIRI$Treat_Type))

sapling_PIRI$Region = factor(sapling_PIRI$Region,
                          levels = unique(sapling_PIRI$Region))

sapling_PIRI$Site = factor(sapling_PIRI$Site,
                        levels = unique(sapling_PIRI$Site))
```
```





















