---
title: "Small seeding models for Tony & Kevin"
author: "Kathleen Stutzman"
date: "2024-01-25"
output: html_document
---

Hidden below is code for importing the data set and log transforming variables. The best model (lowest AIC & using lrtest) is presented below, model fit checked, and pairwise comparison for Treatment Type.  

I did create and test models without treatment type, but they always have higher AIC and worse model fit that those with (probably due to the control category...)  

##### Variables investigated:
- **Treat_Type** - treatment type
- **PIRI** - pitch pine seedling count (1m^2^)
- **SO** - shrub oak seedling count (1m^2^)
- **other** - other seedling count (1m^2^)
- **BA_HA** - basal area per hectare
- **BA_piri** - pitch pine basal area per hectare
- **Mineral** - exposed mineral soil (1m^2^)
- **avgLD** - average leaf litter depth (1m^2^)
- **Veg_Total** - total understory plant cover (1m^2^)
- **TFT** - time between treatment and sampling (used as an offset)
- **Site/Plot_No** - random effect of plots within sites

*Variables are log transformed (l. or _l)*

## Pitch Pine Regeneration

```{r Libraries, message=FALSE, warning=FALSE, include=FALSE}
# libraries -------------------
library(dplyr)
library(tidyverse)
library(ggplot2)
library(GGally)
library(glmmTMB)
library(TMB)
library(emmeans)
library(DHARMa)
library(lmtest)
```

```{r, setup, include=FALSE}
# Set WD:
knitr::opts_knit$set(root.dir = '/Users/user/Desktop/Data/Regen_RProj/')
```

```{r, include=FALSE, message=FALSE, warning=FALSE}
# Import seedling data

source("Scripts/SS_Import.R")

ss_merge2 <- ss_merge %>% 
  select(Plot_No, Region, Treat_Type, Site, Species_Groups, Total) #this is dropping stump sprout, browse, and germinate data

ss_merge2 <- ss_merge2 %>% 
  pivot_wider(names_from = Species_Groups, values_from = Total)

time_since <- read_csv("CleanData/Treat_Year_Data.csv")

ss_merge3 <- merge(ss_merge2, time_since, by = "Site")
#log transform time from treatment data
ss_merge3$l.TFT <- log(ss_merge3$Time_from_Treat)

source("Scripts/Add_BA.R")

# merge with ss dataset -------------------
ss_merge4 <- merge(ss_merge3, prism_BA, by = "Plot_No")

source("Scripts/Ground_Data.R")

# merge with ss dataset -------------------
ss_merge5 <- merge(ss_merge4, ground3, by = "Plot_No")

rm(ss_merge2,
   ss_merge3,
   ss_merge4,
   time_since)

source("Scripts/Veg_Data.R")

# merge with ss dataset 
ss.all <- merge(ss_merge5, veg3, by = "Plot_No")

ss.all$l.PIRI <- log(ss.all$PIRI + 1)
ss.all$l.SO <- log(ss.all$Shrub_Oak + 1)
ss.all$l.other <- log(ss.all$Other + 1)

ss.all2 <- ss.all %>% 
  select(Treat_Type, Region, Site, Plot_No, PIRI, l.PIRI, Shrub_Oak, l.SO, Other, l.other, Time_from_Treat, l.TFT, BA_HA, l.BA_HA, PIRI.BA_HA, l.BA_piri, Mineral_Soil, l.Mineral, Litter_Duff, avgLD, avgLD_l, Veg_Total, l.Veg_Total) %>% 
  arrange(Treat_Type)
```

### Best small seedling pitch pine model
```{r}
ss.m4 <- glmmTMB(PIRI ~ Treat_Type + l.SO + avgLD_l + l.Veg_Total + offset(l.TFT) + (1|Site/Plot_No),
                 data = ss.all2,
                 family = poisson)
summary(ss.m4)
```

### Check model fit
```{r}
ss.m4_sr <- simulateResiduals(ss.m4, n = 1000, plot = TRUE) #looks pretty good

testResiduals(ss.m4_sr) #passes uniformity, dispersion, and outliers
testQuantiles(ss.m4_sr) #passes
testZeroInflation(ss.m4_sr) #passes
```

### Pairwise for treatment type
```{r}
emmeans(ss.m4, specs = pairwise ~ Treat_Type, adjust = 'Tukey', type = 'response')
```

### Graph pitch pine models
```{r, warning=FALSE, message=FALSE, fig.align='center'}
library(sjPlot)
library(sjlabelled)
library(sjmisc)

set_theme(base = theme_classic(),
          theme.font = 'serif',
          axis.title.size = 1.5,
          axis.textsize.x = 1.5,
          axis.textsize.y = 1.5,
          title.size = 2.5,
          title.align = "center",
          legend.pos = "right",
          legend.size = 1.5,
          legend.title.size = 1.5,
          #legend.bordercol = "black",
          legend.item.size = .75)

# SS PIRI plot 1 ------------------- LD vs. TT
plot_model(ss.m4, 
                       type = 'pred', 
                       terms = c('avgLD_l', 'Treat_Type'),
                       axis.title = c("Average Leaf Litter Depth (log transformed)", "Total Count of Pitch Pine"),
                       title = "Predicted Counts of Pitch Pine \n Seedlings <50cm",
           legend.title = "Treatment Type",
           line.size = 1,
           value.offset = 'Treat_Type',
           ci.lvl = NA,
           colors = c("#D8B70A", "#02401B", "#A2A475", "#81A88D", "#972D15")) 


# SS PIRI plot 2 ------------------- Veg vs TT
plot_model(ss.m4, 
                       type = 'pred', 
                       terms = c('l.Veg_Total', 'Treat_Type'),
                       axis.title = c("Average Understory Vegetation Cover (log transformed)", "Total Count of Pitch Pine"),
                       title = "Predicted Counts of Pitch Pine \n Seedlings <50cm",
           legend.title = "Treatment Type",
           line.size = 1,
           value.offset = 'Treat_Type',
           ci.lvl = NA,
           colors = c("#D8B70A", "#02401B", "#A2A475", "#81A88D", "#972D15")) 


# SS PIRI pot 3 ------------------- Shrub Oak seedling vs TT
plot_model(ss.m4, 
                       type = 'pred', 
                       terms = c('l.SO', 'Treat_Type'),
                       axis.title = c("Shrub oak small seedling counts (log transformed)", "Total Count of Pitch Pine"),
                       title = "Predicted Counts of Pitch Pine \n Seedlings <50cm",
           legend.title = "Treatment Type",
           line.size = 1,
           value.offset = 'Treat_Type',
           ci.lvl = NA,
           colors = c("#D8B70A", "#02401B", "#A2A475", "#81A88D", "#972D15")) 



# SS PIRI plot 4 ------------------- Time from Treatment vs TT
plot_model(ss.m4, 
                       type = 'pred', 
                       terms = c('l.TFT', 'Treat_Type'),
                       axis.title = c("Time from Treatment (log transformed)", "Total Count of Pitch Pine"),
                       title = "Predicted Counts of Pitch Pine \n Seedlings <50cm",
           legend.title = "Treatment Type",
           line.size = 1,
           value.offset = 'Treat_Type',
           ci.lvl = NA,
           colors = c("#D8B70A", "#02401B", "#A2A475", "#81A88D", "#972D15")) 
```
  




## Shrub Oak Regeneration
This includes QUIL (scrub oak) & QUPR (dwarf chinkapin oak)

### Best small seedling shrub oak model
```{r}
so.ss11 <- glmmTMB(Shrub_Oak ~ Treat_Type + l.PIRI + offset(l.TFT) + (1|Site/Plot_No),
                 data = ss.all2,
                 family = genpois)
summary(so.ss11)
```

### Check model fit
```{r}
so.ss11_sr <- simulateResiduals(so.ss11, n = 1000, plot = TRUE)
testZeroInflation(so.ss11_sr) #passes
testResiduals(so.ss11_sr) #passes uniformity & dispersion
testQuantiles(so.ss11_sr) #quantiles are not good
```
- From what I've read, models with significant quantile deviations can still be good models / used. There are some distribution issues. I've switched to genpois distro here, which just barely gets the job done. Models using poisson or com-pois distros are underdispersed. *Additional thoughts appreciated*

### Pairwise for treatment type
```{r}
emmeans(so.ss11, specs = pairwise ~ Treat_Type, adjust = 'Tukey', type = 'response')
```
#### MowRX, FallRx, and SpringRx significantly different from Control.


### Graph shrub oak seedling model
```{r, warning=FALSE, message=FALSE, fig.align='center'}
# SS SO plot 1 ------------------- PIRI vs. TT
plot_model(so.ss11, 
                       type = 'pred', 
                       terms = c('l.PIRI', 'Treat_Type'),
                       axis.title = c("Pitch pine small seedling counts (log transformed)", "Total Count of Shrub Oak"),
                       title = "Predicted Counts of Shrub Oak \n Seedlings <50cm",
           legend.title = "Treatment Type",
           line.size = 1,
           value.offset = 'Treat_Type',
           ci.lvl = NA,
           colors = c("#D8B70A", "#02401B", "#A2A475", "#81A88D", "#972D15")) 
````






















