---
title: "Small seeding models for Tony & Kevin"
author: "Kathleen Stutzman"
date: "2024-01-25"
output: html_document
---

Hidden below is code for importing the data set and log transforming variables. The best model (lowest AIC & using lrtest) is presented below, model fit checked, and pairwise comparison for Treatment Type.  

I did create and test models without treatment type, but they always have higher AIC and worse model fit that those with (probably due to the control category...)  

##### Variables investigated:
- **Treat_Type** - treatment type
- **PIRI** - pitch pine seedling count (1m^2^)
- **SO** - shrub oak seedling count (1m^2^)
- **other** - other seedling count (1m^2^)
- **BA_HA** - basal area per hectare
- **BA_piri** - pitch pine basal area per hectare
- **Mineral** - exposed mineral soil (1m^2^)
- **avgLD** - average leaf litter depth (1m^2^)
- **Veg_Total** - total understory plant cover (1m^2^)
- **TFT** - time between treatment and sampling (used as an offset)
- **Site/Plot_No** - random effect of plots within sites

*Variables are log transformed (l. or _l)*

## Pitch Pine Regeneration

```{r Libraries, message=FALSE, warning=FALSE, include=FALSE}
# libraries -------------------
library(dplyr)
library(tidyverse)
library(ggplot2)
library(GGally)
library(glmmTMB)
library(TMB)
library(emmeans)
library(DHARMa)
library(lmtest)
library(ggeffects)
```

```{r, setup, include=FALSE}
# Set WD:
knitr::opts_knit$set(root.dir = '/Users/user/Desktop/Data/Regen_RProj/')
```

```{r, include=FALSE, message=FALSE}
# Import seedling data

source("Scripts/SS_Import.R")

ss_merge2 <- ss_merge %>% 
  select(Plot_No, Region, Treat_Type, Site, Species_Groups, Total) #this is dropping stump sprout, browse, and germinate data

ss_merge2 <- ss_merge2 %>% 
  pivot_wider(names_from = Species_Groups, values_from = Total)

time_since <- read_csv("CleanData/Treat_Year_Data.csv")

ss_merge3 <- merge(ss_merge2, time_since, by = "Site")
#log transform time from treatment data
ss_merge3$l.TFT <- log(ss_merge3$Time_from_Treat)

source("Scripts/Add_BA.R")

# merge with ss dataset -------------------
ss_merge4 <- merge(ss_merge3, prism_BA, by = "Plot_No")

source("Scripts/Ground_Data.R")

# merge with ss dataset -------------------
ss_merge5 <- merge(ss_merge4, ground3, by = "Plot_No")

rm(ss_merge2,
   ss_merge3,
   ss_merge4,
   time_since)

source("Scripts/Veg_Data.R")

# merge with ss dataset 
ss.all <- merge(ss_merge5, veg3, by = "Plot_No")

ss.all$l.PIRI <- log(ss.all$PIRI + 1)
ss.all$l.SO <- log(ss.all$Shrub_Oak + 1)
ss.all$l.other <- log(ss.all$Other + 1)

ss.all2 <- ss.all %>% 
  select(Treat_Type, Region, Site, Plot_No, PIRI, l.PIRI, Shrub_Oak, l.SO, Other, l.other, Time_from_Treat, l.TFT, BA_HA, l.BA_HA, PIRI.BA_HA, l.BA_piri, Mineral_Soil, l.Mineral, Litter_Duff, avgLD, avgLD_l, Veg_Total, l.Veg_Total) %>% 
  arrange(Treat_Type)
```

### Best small seedling pitch pine model
```{r}
ss.m4 <- glmmTMB(PIRI ~ Treat_Type + l.SO + avgLD_l + l.Veg_Total + offset(l.TFT) + (1|Site/Plot_No),
                 data = ss.all2,
                 family = poisson)
summary(ss.m4)
```

### Check model fit
```{r}
ss.m4_sr <- simulateResiduals(ss.m4, n = 1000, plot = TRUE) #looks pretty good

testResiduals(ss.m4_sr) #passes uniformity, dispersion, and outliers
testQuantiles(ss.m4_sr) #passes
testZeroInflation(ss.m4_sr) #passes
```

### Pairwise for treatment type
```{r}
emmeans(ss.m4, specs = pairwise ~ Treat_Type, adjust = 'Tukey', type = 'response')
```
#### FallRx, Harvest, MowRx, SpringRx all significantly different from Control - no treatment significantly different from another 


### Graph pitch pine models
```{r, fig.align='center', echo=FALSE}
ss.p1 <- ggpredict(ss.m4, terms = c("avgLD_l", "Treat_Type"))


ggplot(ss.p1, aes(x, predicted, color = group))+
  geom_line(linewidth = 1.25,  show.legend = FALSE) +
  scale_color_manual(values = c("#D8B70A",  "#02401B", "#A2A475", "#81A88D", "#972D15"))+
  theme_classic()+
  theme(panel.background = element_blank()) +
  theme(axis.text = element_text(size = 14), axis.title = element_text(size = 20))+
  labs(x = "Average leaf litter depth (log transformed)",
       y = "Pitch pine stems \n (corrected for time from treatment)")

# PIRI SS plot 2 -------------------
ss.p2 <- ggpredict(ss.m4, terms = c("l.Veg_Total", "Treat_Type"))


ggplot(ss.p2, aes(x, predicted, color = group))+
  geom_line(linewidth = 1.25,  show.legend = FALSE) +
  scale_color_manual(values = c("#D8B70A",  "#02401B", "#A2A475", "#81A88D", "#972D15"))+
  theme_classic()+
  theme(panel.background = element_blank()) +
  theme(axis.text = element_text(size = 14), axis.title = element_text(size = 20))+
  labs(x = "Average understory vegetation cover \n (log transformed)",
       y = "Pitch pine stems \n (corrected for time from treatment)")


# PIRI SS plot 3 -------------------
ss.p3 <- ggpredict(ss.m4, terms = c("l.SO", "Treat_Type"))


ggplot(ss.p3, aes(x, predicted, color = group))+
  geom_line(linewidth = 1.25) +
  scale_color_manual(values = c("#D8B70A",  "#02401B", "#A2A475", "#81A88D", "#972D15"))+
  theme_classic()+
  theme(panel.background = element_blank()) +
  theme(axis.text = element_text(size = 14), axis.title = element_text(size = 20),
        legend.position = "right", legend.text = element_text(size = 18), legend.title = element_text(size = 22))+
  labs(x = "Shrub oak seedling counts (log transformed)",
       y = "Pitch pine stems \n (corrected for time from treatment)",
       color = "Treatment Type")
```




## Shrub Oak Regeneration
This includes QUIL (scrub oak) & QUPR (dwarf chinkapin oak)

### Best small seedling shrub oak model
```{r}
so.ss8a <- glmmTMB(Shrub_Oak ~ Treat_Type + offset(l.TFT) + (1|Site/Plot_No),
                 data = ss.all2,
                 ziformula = ~Region,
                 family = nbinom1)
summary(so.ss8a)
```

### Check model fit
```{r}
so.ss8a_sr <- simulateResiduals(so.ss8a, n = 1000, plot = TRUE)
testResiduals(so.ss8a_sr) #passes
testQuantiles(so.ss8a_sr) # p = 0.000006; this is the best so far
```
- From what I've read, models with significant quantile deviations can still be good models / used.

### Pairwise for treatment type
```{r}
emmeans(so.ss8a, specs = pairwise ~ Treat_Type, adjust = 'Tukey', type = 'response')
```
#### MowRX, FallRx, and SpringRx significantly different from Control.


### Graph shrub oak seedling model
```{r, fig.align='center', echo=FALSE}
so.p1 <- ggpredict(so.ss8a, terms = c("l.TFT", "Treat_Type"))

ggplot(so.p1, aes(x, predicted, color = group))+
  geom_boxplot() +
  scale_color_manual(values = c("#D8B70A",  "#02401B", "#A2A475", "#81A88D", "#972D15"))+
  theme_classic()+
  theme(panel.background = element_blank()) +
  theme(axis.text = element_text(size = 14), axis.title = element_text(size = 20), legend.position = "right", legend.text = element_text(size = 14), legend.title = element_text(size = 18)) +
  labs(x=NULL,
       y = "Shrub Oak Small Seedlings \n (corrected for time from treatment)",
       color = "Treatment Type")


ggplot(so.p1, aes(x, predicted, color = group))+
  geom_line(linewidth = 1.25) +
  scale_color_manual(values = c("#D8B70A",  "#02401B", "#A2A475", "#81A88D", "#972D15"))+
  theme_classic()+
  theme(panel.background = element_blank()) +
  theme(axis.text = element_text(size = 14), axis.title = element_text(size = 20), legend.position = "right", legend.text = element_text(size = 14), legend.title = element_text(size = 18)) +
  labs(x = "Time from Treatment (log transformed)",
       y = "Shrub Oak Stems \n (corrected for time from treatment)",
       color = "Treatment Type")

# this should really be a boxplot?, not sure how this should be visulaized, as TFT isn't really a variable but a factor in the model

```





















