---
title: "Understory veg ordination for MA & LI"
author: "Kathleen Stutzman"
date: "2024-02-26"
output: html_document
---

Libraries:
```{r, message=FALSE, warning=FALSE}
# libraries -------------------
library(ggplot2)
library(vegan)
library(ggvegan)
library(tidyverse)
```

Set working directory
```{r, setup}
knitr::opts_knit$set(root.dir = '/Users/user/Desktop/Data/Regen_RProj/')
```

Functions:
```{r}
# functions -------------------
source("Functions/veganCovEllipse.R")
```


#### Source, tranform, select, and relativize MA & LI veg data (hidden)
```{r, include=FALSE}
source("Scripts/Veg_Data.R")
rm(veg3)

#split overall data from veg cover
meta_mali <- veg2 %>%
  filter(Region == c("MA", "LI")) %>% 
  select(Region, Treat_Type, Site, Plot_No) %>% 
  distinct(Plot_No, .keep_all = TRUE)

#pivot veg data wider and fill in zeros
mali_wider <- veg2 %>% 
  filter(Region == c("MA", "LI")) %>% 
  select(Plot_No, Species_Code, SP_Cover_midpoint) %>% 
  group_by(Plot_No) %>% 
  pivot_wider(names_from = Species_Code, values_from = SP_Cover_midpoint, values_fill = 0) 

#create dataframe of site and plot
mali_site <- veg2 %>% 
  filter(Region == c("MA", "LI")) %>% 
  select(Site, Plot_No) %>% 
  group_by(Site) %>% 
  distinct(Plot_No, .keep_all = TRUE)

# merge site and veg data; remove plot number
mer.mali <- merge(mali_site, mali_wider, by = "Plot_No") %>% 
  select(-Plot_No)

#THIS IS 25 plots or more
mali_5PERC <- mali_wider %>% 
  select(Plot_No,
         QUIL,
         GAPR,
         VAPA,
         CAPE,
         GABA,
         VAAN,
         PTAQ,
         RUSP,
         KAAN,
         PIRI,
         ACRU,
         #LYQU,
         QUCO,
         MELI,
         #CEOR,
         COPE,
         PRSE,
         QUPR,
         #PAQU,
         #CEAM,
         QUAL,
         PIST,
         #ARME,
         #LYBO,
         POUN11,
         #RHGL
         )

# add site and lose plot number -------------------
mali5 <- merge(mali_5PERC, mali_site) %>% 
  select(-Plot_No) %>% 
  select(Site, everything())

mali5_avg <- mali5 %>% 
  group_by(Site) %>% 
  summarise_at(vars(QUIL:POUN11), mean) %>% 
  column_to_rownames(var = "Site")

rm(mali5, mali_5PERC, mer.mali)

# relativize data, dividing each cell by the column total (to bring common and rare species closer together) -------------------
rel_mali5 <- decostand(mali5_avg, method = 'total', MARGIN = 2)
```


### What's different about 5% min plot species for this region?
```{r}
mali_wider_log <- as.data.frame(lapply(mali_wider, as.logical))

ma.li <- as.data.frame(colSums(mali_wider_log)) %>% 
  rownames_to_column(var = "Species")

# Species from the other 5% calcs
# In NH/ME - no LYQU, CEOR, PAQU, CEAM, ARME, LYBO, RHGL

# For this data set , 5% is ~11 plots:
# Would add: SMGL
# If I stuck to at least 11, would lose: ACRU, COPE, PRSE, PIST, POUN11

rm(mali_wider_log, ma.li)
```

#### Source and transform envr & categorical data; merge for predictor df to compare nmds against (hidden)
```{r, include=FALSE}
time_since <- read_csv("CleanData/Treat_Year_Data.csv") # this is a dataframe with one row for each site

# make sure ggpubs not on, or this won't load
source("Scripts/Add_BA.R") #work on prism_BA, has total BA figures - do I need to change these as they were collected on a different scale?

ma.prism1 <- prism_BA %>% 
  select(Plot_No, BA_HA, PIRI.BA_HA)

ma.prism2 <- merge(ma.prism1, mali_site, by = "Plot_No") %>% 
  select(-Plot_No) %>% 
  select(Site, everything())

ma.prism4veg <- ma.prism2 %>% 
  group_by(Site) %>% 
  summarise_at(vars(BA_HA:PIRI.BA_HA), mean)

rm(prism_BA, ma.prism1, ma.prism2)


source("Scripts/Ground_Data.R") #work on ground 3

ma.ground4 <- ground3 %>% 
  select(Plot_No, Lichen, Moss, Rock, Mineral_Soil, Wood, Litter_Duff, avgLD)

ma.ground5 <- merge(ma.ground4, mali_site, by = "Plot_No") %>% 
  select(Site, everything()) %>% 
  select(-c(Plot_No))

ma.ground4veg <- ma.ground5 %>% 
  group_by(Site) %>% 
  summarise_at(vars(Lichen:avgLD), mean)

rm(ground3, ma.ground4, ground)

ma2 <- meta_mali %>% 
  select(Region, Treat_Type, Site) %>% 
  distinct(Site, .keep_all = TRUE)

mam1 <- merge(ma2, time_since, by = "Site")

mam2 <- merge(mam1, ma.ground4veg, by = "Site")

ma.meta.data_veg <- merge(mam2, ma.prism4veg, by = "Site") %>% 
  select(-c(Inventory_Yr, Treat_Yr))


rm(mam2, mam1, time_since, ma.ground4veg, ma.prism4veg)

rm(prism, veg2)
```

### Run NMDS
```{r}
# 3 dimensions
mali_3d <- metaMDS(rel_mali5,
                   distance = 'bray',
                   k = 3,
                   trymax = 20,
                   autotransform = FALSE)
# converges, stress of 0.13 - 0.15

# 2 dimensions
mali_2d <- metaMDS(rel_mali5, 
                   distance = 'bray', 
                   k = 2,
                   try = 40,
                   trymax = 40,
                   autotransform = FALSE)
# stress is too high to represent in 2 dimensions
```


### Make correlation data for bi-plots
```{r}
ma.veg_enr <- envfit(mali_3d, ma.meta.data_veg)
ma.veg_spc.envr <- envfit(mali_3d, rel_mali5)

# site data -------------------
ma.site.scrs <- as.data.frame(scores(mali_3d, display = "sites"))

ma.site.scrs <- cbind(ma.site.scrs, Treat_Type = ma.meta.data_veg$Treat_Type)
ma.site.scrs <- cbind(ma.site.scrs, Region = ma.meta.data_veg$Region)

# species data   -------------------
ma.spp.scrs <- as.data.frame(scores(ma.veg_spc.envr, display = "vectors"))
ma.spp.scrs <- cbind(ma.spp.scrs, Species = rownames(ma.spp.scrs))
ma.spp.scrs <- cbind(ma.spp.scrs, pval = ma.veg_spc.envr$vectors$pvals)

sig.ma_spp.scrs <- subset(ma.spp.scrs, pval<=0.05)

# envr data -------------------
ma.envr.scrs <- as.data.frame(scores(ma.veg_enr, display = "vectors"))
ma.envr.scrs <- cbind(ma.envr.scrs, env.variables = rownames(ma.envr.scrs))
ma.envr.scrs <- cbind(ma.envr.scrs, pval = ma.veg_enr$vectors$pvals)
ma.sig_envr.scrs <- subset(ma.envr.scrs, pval<=0.05)
```


## Need to think about how to represent 3D



### Run PerMANOVA to look at the impacts of region and treatment on community composition
```{r}
summary(as.factor(ma.meta.data_veg$Treat_Type))
```

```{r}
ma.tt <- adonis2(rel_mali5 ~ Treat_Type*Region, data = ma.meta.data_veg)

print(ma.tt)

#each are significant, but the interaction is not
```























