---
title: "Veg descriptive statistics by site"
author: "Kathleen Stutzman"
date: "2024-02-20"
output: html_document
---

This script calculates species richness, species diversity (Shannon Index), species numbers, evenness, and beta diversity for veg data


Libraries:
```{r Libraries, message=FALSE, warning=FALSE}
# libraries -------------------
library(vegan)
library(tidyverse)
library(plyr)
library(dplyr)
library(adespatial)
```

Set working directory
```{r, setup}
knitr::opts_knit$set(root.dir = '/Users/user/Desktop/Data/Regen_RProj/')
```

Source functions:
```{r}
# source functions -------------------
source("Functions/species_richness.R")
source("Functions/shannon_index.R")
source("Functions/spec_num.R")
source("Functions/Pielous_J.R")
```

Source veg data:
```{r}
#start by sourcing the veg script
source("Scripts/Veg_Data.R")

rm(veg3)#this is used for total cover in GLMMs; not used in this script and doesn't import correctly due to use of dplyr package
```

Split overall data from veg cover; pivot wider to fill in missing 0s; create df of site and plot to reattach later
```{r}
#split overall data from veg cover
veg_meta <- veg2 %>% 
  select(Region, Treat_Type, Site, Plot_No) %>% 
  distinct(Plot_No, .keep_all = TRUE)

#pivot veg data wider and fill in zeros
veg_wider <- veg2 %>% 
  select(Plot_No, Species_Code, SP_Cover_midpoint) %>% 
  group_by(Plot_No) %>% 
  pivot_wider(names_from = Species_Code, values_from = SP_Cover_midpoint, values_fill = 0) 

#create dataframe of site and plot
veg_site <- veg2 %>% 
  select(Site, Plot_No) %>% 
  group_by(Site) %>% 
  distinct(Plot_No, .keep_all = TRUE)

#reattach site info and reorder
veg_merge1 <- merge(veg_site, veg_wider, by = "Plot_No") %>% 
   select(Site, Plot_No, everything())
```

Create df without plot numbers; then split df by site
```{r}
#remove plot number
veg_merge3 <- veg_merge1 %>% 
  select(-Plot_No)

#split df by site
split_veg <- split(veg_merge3[-1], veg_merge3$Site)
```

##### Species Richness
```{r}
# Species Richness (jackknife and chao etc estimates) -------------------
sp <- as.data.frame(sapply(split_veg, species_richness))

#transpose
spcrich_bysite <- as.data.frame(t(sp))

rm(sp)
```

##### Diversity (Shannon-Wiener Index)
```{r}
# Shannon index -------------------
si <- as.data.frame(sapply(split_veg, shannon_index))

#transpose
si_bysite <- as.data.frame(t(si))

#name columns
colnames(si_bysite) <- c("S.I._Mean", "S.I._SD")

rm(si)
```

##### Species Number
```{r}
# Number of species -------------------
spc_n <- as.data.frame(sapply(split_veg, spec_num))

#transpose
spcnum_bysite <- as.data.frame(t(spc_n))

#name columns
colnames(spcnum_bysite) <- c("Spc.Num_Mean", "Spc.Num_SD")

rm(spc_n)
```

##### Evenness (Pielou's J)
This is the SW Index / log (species number). For three sites (one each in APB_Hippo, RPD2_B2, & RPD2_B7), they have a species number of 1. The log of 1 is 0 and anything divided by 0 is NaN. Therefore, these three subplots are dropped from the calculations

```{r}
even1 <- as.data.frame(sapply(split_veg, Pielous_J))

even2 <- as.data.frame(t(even1))

colnames(even2) <- c("Pielous_J_mean", "Pielous_J_SD")

rm(even1)
```

Hidden below is code to look at the three subplots, if desired
```{r, include=FALSE, eval=FALSE}
RPD2_B7
HIPPO <-  veg_merge3 %>% 
  filter(Site == "RPD2_B7") %>% 
  select(-Site)

H <- diversity(HIPPO)
S <- specnumber(HIPPO)
J <- H/(log(S))

print(J)

rm(H, S, J, HIPPO)
#just switch out site name
```

### Sorensen dissimilarity index
```{r}
# creating a dataframe where observations from every subplot within a site are added together to use for P/A calculations in sorensen
veg4soren <- veg_merge3 %>% 
  group_by(Site) %>% 
  summarise_at(vars(QUCO:HECA), sum) %>% 
  column_to_rownames(var = "Site")

# creating df of row id and site id to reattach to beta diversity metrics later
site_rowid <- veg_merge3 %>% 
  group_by(Site) %>% 
  summarise_at(vars(QUCO:HECA), sum) %>% 
  rowid_to_column(var="rowid") %>% 
  select(Site, rowid)
```

Calculate beta diversity
```{r}
BDiv <- beta.div.comp(veg4soren, coef = "S", quant = FALSE)
#overall values
BetaD.df <- as.data.frame(BDiv$part)
#distance matrix for all sites
BetaD.distances <- as.matrix(BDiv$D)
#replacement values matrix for all sites
BetaD.repl.matrix <- as.matrix(BDiv$repl)
#richness values matrix for all sites
BetaD.rich.matrix <- as.matrix(BDiv$rich)
```

#### Overall beta diversity metrics
```{r}
print(BetaD.df)
```

#### Distance matrix for all sites
```{r}
print(BetaD.distances)
```



Local contributions to beta diversity; partitioned into richness and replacement
```{r}
# replacement
replacement <- LCBD.comp(BDiv$repl, sqrt.D = TRUE)

rep.df <- as.data.frame(replacement$LCBD)

names(rep.df)[names(rep.df) == "replacement$LCBD"] <- "replace.lcbd"

rep.df <-  rep.df %>% 
  rowid_to_column(var="rowid")

rep.df2 <- merge(rep.df, site_rowid, by = "rowid") %>%
  select(Site, replace.lcbd) %>% 
  column_to_rownames(var = "Site")


#richness
richness <- LCBD.comp(BDiv$rich, sqrt.D = TRUE)

rich.df <- as.data.frame(richness$LCBD)

names(rich.df)[names(rich.df) == "richness$LCBD"] <- "richness.lcbd"

rich.df <- rich.df %>% 
  rowid_to_column(var = "rowid")

rich.df2 <- merge(rich.df, site_rowid, by = "rowid") %>% 
  select(Site, richness.lcbd) %>% 
  column_to_rownames(var = "Site")

rm(rep.df, rich.df)
```

### Merge DS together
```{r}
# merging data sets of DS
m1 <- merge(spcrich_bysite, si_bysite, by="row.names") %>% 
  column_to_rownames(var = "Row.names")


m2 <- merge(m1, spcnum_bysite, by="row.names") %>% 
  column_to_rownames(var = "Row.names")

m3 <- merge(m2, even2, by = "row.names") %>% 
  column_to_rownames(var = "Row.names")


m4 <- merge(m3, rich.df2, by = "row.names") %>% 
  column_to_rownames(var = "Row.names")

veg_sd <- merge(m4, rep.df2, by = "row.names") %>% 
  column_to_rownames(var = "Row.names")


rm(m1, 
   m2, 
   m3, 
   m4)
```


```{r}
print(veg_sd)
```


































