---
title: "Sapling PIRI GLMM"
author: "Kathleen Stutzman"
date: "2024-01-04"
output: html_document
---

## Introduction:
This script, originally written as an R script, will be much more readable as an R Markdown. Therefore, I am translating it into a markdown file.

# Pitch Pine Sapling GLMM code

Libraries:
```{r Libraries, message=FALSE, warning=FALSE}
# libraries -------------------
library(dplyr)
library(tidyverse)
library(ggplot2)
library(lme4)
library(lmerTest)
library(glmmTMB)
library(TMB)
library(emmeans)
library(DHARMa)
```

Set working directory
```{r, setup}
knitr::opts_knit$set(root.dir = '/Users/user/Desktop/Data/Regen_RProj/')
```

Functions:
```{r}
# functions -------------------

```

Global variables:
```{r}
file_folder_SA <- "CleanData/sapling_CSV/"

options(scipen = 999) #to display actual values

file_names_SA <- list.files(path=file_folder_SA)
```

For loop to read in files:
```{r}
for (i in seq_along(file_names_SA)) {
  assign(paste0(file_names_SA[i], i),
         read.csv(paste0(file_folder_SA, file_names_SA[i])))
}

# Adding treatment types -------------------
SA_Control_Data.csv1$Treat_Type <- "Control"
SA_Fall_Burn_Data.csv2$Treat_Type <- "FallRx"
SA_Mow_Burn_Data.csv3$Treat_Type <- "MowRx"
SA_Spring_Burn_Data.csv4$Treat_Type <- "SpringRx"
SA_Thinned_Data.csv5$Treat_Type <- "Harvest"

# merge into one data set -------------------
sapling <- rbind(SA_Control_Data.csv1, 
                 SA_Fall_Burn_Data.csv2, 
                 SA_Mow_Burn_Data.csv3, 
                 SA_Spring_Burn_Data.csv4, 
                 SA_Thinned_Data.csv5) 


# remove data sets to keep environment clean-------------------
rm(SA_Control_Data.csv1,
   SA_Fall_Burn_Data.csv2,
   SA_Mow_Burn_Data.csv3,
   SA_Spring_Burn_Data.csv4,
   SA_Thinned_Data.csv5)
```

Creating groupings of species: PIRI, Shrub Oak, & Other:
```{r}
sapling$Species_Groups <- NA

# PIRI
sapling$Species_Groups <- ifelse(sapling$Species_Code == 'PIRI', 'PIRI', sapling$Species_Groups)

# Shrub Oak
sapling$Species_Groups <- ifelse(sapling$Species_Code %in% c('QUIL','QUPR'), 'Shrub Oak', sapling$Species_Groups)

#Other  
sapling$Species_Groups <- ifelse(sapling$Species_Code %in% c('ACRU', 'BEPO', 'QUCO', 'QUAL', 'QUVE', 'QUST', 'QURU', 'QUPR', 'PRSE', 'FRPE', 'PRVI', 'PODE', 'POGR', 'SAAL', 'AMSP', 'FAGR', 'NYSY', 'ACPE', 'ACNE', 'MOSP', 'POTR', 'PRSP', 'PIST', 'ABBA', 'FRAL', 'ROPS', 'RHCA', 'ELUM', 'BELE'), 'Other', sapling$Species_Groups)


# now to check all have been assigned -------------------
check_sa <- sapling %>% 
  filter(if_any(Species_Groups, is.na))

# there are 724 plots that do not have saplings recorded -------------------

rm(check_sa)
```

Manipulate data to create a complete dataset, where implied zeros are now express and combined species groups totals are counted as one:
```{r}
sapling2 <- sapling %>%
  select(Region,
         Treat_Type,
         Site,
         Plot_No,
         Species_Groups, 
         DBH.cm.)

# we should have just tallied amounts instead of recording DBH, so now to convert to tally
sapling2$Tally <- NA

sapling2$Tally <- ifelse(sapling2$DBH.cm. >= 2.5, 1, sapling2$Tally)

sapling2$Tally <- ifelse(sapling2$DBH.cm. == 0, 0, sapling2$Tally)

# check to see all have been assigned -------------------
check_sapling <- sapling2 %>% 
  filter(if_any(Tally, is.na))

rm(check_sapling)

# add duplicates by plot and filter it down -------------------
sapling3 <- sapling2 %>% 
  select(Region,
         Treat_Type,
         Site,
         Plot_No,
         Species_Groups,
         Tally)

#check for correct number of plots (1001)
n_distinct(sapling3$Plot_No)

# split data set to use complete function
sapling_meta <- sapling3 %>% 
  select(Region, Treat_Type, Site, Plot_No) %>% 
  distinct(Plot_No, .keep_all = TRUE)

sapling_plot <- sapling3 %>% 
  select(Plot_No, Species_Groups, Tally)

#duplicates of "other need to be combined
sapling_plot1 <- sapling_plot %>% 
  group_by(Plot_No, Species_Groups) %>% 
  summarise(Total_Tally = sum(Tally)) %>% 
  ungroup()

sapling_plot2 <- sapling_plot1 %>% 
  complete(Plot_No, Species_Groups, fill = list(Total_Tally = 0))

sapling_NA <- sapling_plot2[complete.cases(sapling_plot2),]

sapling_merge <- merge(sapling_meta, sapling_NA, by= "Plot_No")

# now remove datasets no longer needed to keep envr clean -------------------
rm(sapling_meta,
   sapling_plot,
   sapling_plot1,
   sapling_plot2,
   sapling2,
   sapling3,
   sapling_NA)
```

Now looking just at PIRI:
```{r}
# filter for just piri -------------------
sapling_PIRI <- sapling_merge %>% 
  filter(Species_Groups == "PIRI") %>% 
  arrange(Treat_Type)
# PIRI saplings are only in 50 of the 1001 plots and mostly in control sites....


# now to make the explanatory variables factors
sapling_PIRI$Treat_Type = factor(sapling_PIRI$Treat_Type,
                              levels = unique(sapling_PIRI$Treat_Type))

sapling_PIRI$Region = factor(sapling_PIRI$Region,
                          levels = unique(sapling_PIRI$Region))

sapling_PIRI$Site = factor(sapling_PIRI$Site,
                        levels = unique(sapling_PIRI$Site))
```

Import time since data and add it to the PIRI dataset
```{r}
time_since <- read_csv("CleanData/Treat_Year_Data.csv")

sapling_PIRI2 <- merge(sapling_PIRI, time_since, by = 'Site')
#log transform time from treatment data
sapling_PIRI2$l.TFT <- log(sapling_PIRI2$Time_from_Treat)
```

Run the 'Add_BA' script and merge with dataset:
```{r}
source("Scripts/Add_BA.R")

# merge with ls dataset -------------------
sa.merge1 <- merge(sapling_PIRI2, prism_BA, by = 'Plot_No')
```

Run 'Ground_Data.R' script and add it to PIRI dataset:
```{r}
source("Scripts/Ground_Data.R")

# merge with ls dataset -------------------
sa_alldata <- merge(sa.merge1, ground3, by = 'Plot_No')

rm(#sa.merge1, keeping this data set to be able to run models that operate on all the observations, as LD not important
   sapling_PIRI2)
```

Hidden below are models with convergence problems:
```{r, eval=FALSE, include=FALSE}
sa.nb_all <- glmer.nb(Total_Tally~Treat_Type + l.BA_HA + l.BA_piri + l.Mineral + avgLD_l + (1|Site/Plot_No) + offset(l.TFT),
                      data = sa_alldata)
summary(sa.nb_all) #failed to converge

sa.nb2 <- glmer.nb(Total_Tally~Treat_Type + l.Mineral + avgLD_l + (1|Site/Plot_No) + offset(l.TFT),
                      data = sa_alldata)
summary(sa.nb2) #failed to converge, AIC is 217

rm(sa.nb2, sa.nb_all)
```

Models that converge:
```{r}
sa.p3 <- glmer(Total_Tally~Treat_Type + avgLD_l + (1|Site/Plot_No) + offset(l.TFT),
               data = sa_alldata,
               family = poisson)
summary(sa.p3) #first model that didn't fail to converge, AIC is 213.6

######## this is the best model
sa.p4 <- glmer(Total_Tally~Treat_Type + (1|Site/Plot_No) + offset(l.TFT),
               data = sa_alldata,
               family = poisson)
summary(sa.p4) #AIC is 211.7

# would like to try this model again, but with all the data (as model is not using veg/soil data, which cuts sample size in half)
sa.p5 <- glmer(Total_Tally~Treat_Type + (1|Site/Plot_No) + offset(l.TFT),
               data = sa.merge1,
               family = poisson)
summary(sa.p5) #it fails to converge with all data
rm(sa.pa5)

sa.p6 <- glmer(Total_Tally~Treat_Type + (1|Site/Plot_No),
               data = sa_alldata,
               family = poisson)
summary(sa.p6) #AIC is 211.3

sa.p7 <- glmer(Total_Tally~Treat_Type + l.BA_piri + (1|Site/Plot_No) + offset(l.TFT),
               data = sa_alldata,
               family = poisson)
summary(sa.p7)
```

Test models:
```{r}
library(lmtest)
lrtest(sa.p4, sa.p6) #same p-value as the offset in other tests, wonder what that's about

lrtest(sa.p4, sa.p7) #p = .2, not worth the l.BA_piri

rm(sa.p3, sa.p6, sa.p7)
```

Run DHARMa package to check model fit:
```{r}
sa.p4_sr <- simulateResiduals(fittedModel = sa.p4, n=1000, plot = TRUE)

testDispersion(sa.p4_sr)
testZeroInflation(sa.p4_sr)

#no good, model is zero inflated
```

Data is zero inflated. Need to try zero inflated models:
```{r}
######## this is the best model
sa.zi4 <- glmmTMB(Total_Tally~Treat_Type + (1|Site/Plot_No) + offset(l.TFT),
               data = sa_alldata,
               ziformula=~1,
               family = poisson)
summary(sa.zi4) #AIC is 257.7 - this passes the DHARMa test

sa.zi4_sr <- simulateResiduals(fittedModel = sa.zi4, n = 1000, plot = TRUE)

testDispersion(sa.zi4_sr)# PASSES
testOutliers(sa.zi4_sr)
testZeroInflation(sa.zi4_sr) #looking a lot better

plotResiduals(sa.zi4_sr, sa_alldata$Treat_Type, xlab = "Treatment Type", main=NULL)  

#plotResiduals(sa.zi4_sr, sa_alldata$Site, xlab = "Site", main=NULL)

# would like to try this model again, but with all the data
sa.zi9 <- glmmTMB(Total_Tally~Treat_Type + (1|Site/Plot_No) + offset(l.TFT),
                  data = sa.merge1,
                  ziformula=~1,
                  family = poisson)
summary(sa.zi9) #AIC is 411.1 - lets check with dharma 

sa.zi9_sr <- simulateResiduals(fittedModel = sa.zi9, n = 1000, plot = TRUE) #significant differences

testDispersion(sa.zi9_sr)# does not pass
testOutliers(sa.zi9_sr)
testZeroInflation(sa.zi9_sr) #does not pass
#looks like it is better to use half the data, which is what i'm using for the other two measurements, so at least it is consistent

rm(sa.zi9,
   sa.zi9_sr)
```


Here are zero-inflated models that converge but do not pass DHARMa tests
```{r, eval=FALSE}
sa.zi5 <- glmmTMB(Total_Tally~Treat_Type + l.BA_HA + (1|Site/Plot_No) + offset(l.TFT),
                  data = sa_alldata,
                  ziformula=~1,
                  family = poisson)
summary(sa.zi5) #AIC is 215.4, SR don't agree - not better!


sa.zi6 <- glmmTMB(Total_Tally~Treat_Type + l.BA_piri + (1|Site/Plot_No) + offset(l.TFT),
                  data = sa_alldata,
                  ziformula=~1,
                  family = poisson)
summary(sa.zi6) #AIC is 214.2, SR don't agree - not better!

sa.zi7 <- glmmTMB(Total_Tally~Treat_Type + l.BA_piri + (1|Site/Plot_No),
                  data = sa_alldata,
                  ziformula=~1,
                  family = poisson)
summary(sa.zi7) #AIC is 213.9, SR don't agree - not better!

sa.zi8 <- glmmTMB(Total_Tally~Treat_Type + (1|Site/Plot_No),
                  data = sa_alldata,
                  ziformula=~1,
                  family = poisson)
summary(sa.zi8) #doesn't pass the DHARMa test

rm(sa.zi5,
   sa.zi6,
   sa.zi7,
   sa.zi8)
```

Use emmeans for pairwise comparison of Treatment Types:
```{r}
emmeans(sa.zi4, specs = pairwise ~ Treat_Type, adjust = 'Tukey', type = 'response')
#none significant
```

Below is script to create plots. Do not run unless desired. The added packages cause issues for running general GLMM scripts (issues with tidyverse) and therefore are only worth activating if you want to produce graphs or have already loaded all desired analysis
```{r, eval=FALSE}
library(sjPlot)
library(sjlabelled)
library(sjmisc)

set_theme(base = theme_classic(),
          theme.font = 'serif',
          axis.title.size = 1.5,
          axis.textsize.x = 1.5,
          axis.textsize.y = 1.5,
          title.size = 2.5,
          title.align = "center",
          legend.pos = "right",
          legend.size = 1.5,
          legend.title.size = 1.5,
          #legend.bordercol = "black",
          legend.item.size = .75)

plot_model(sa.zi4, 
           type = "pred",
           terms = "Treat_Type")

plot_model(sa.zi4, 
           type = "pred",
           terms = c("l.TFT", "Treat_Type"),
           axis.title = c("Time from Last Treatment (log transformed)", "Total Count of Pitch Pine"),
           title = "Predicted Counts of Pitch Pine Saplings \n >/= 2.5 cm & < 10 cm DBH",
           legend.title = "Treatment Type",
           line.size = 1,
           show.zeroinf = T,
           ci.lvl = NA,
           colors = c("#D8B70A", "#02401B", "#A2A475", "#81A88D", "#972D15"))
```


# Other hidden analysis: early models
```{r, include=FALSE, eval=FALSE}
##################### this is my original analysis of all plot before I added more data / variables to the model -------------------
sa.nb_null <- glmer.nb(Total_Tally~Treat_Type + (1|Site/Plot_No),
                       data = sapling_PIRI)
summary(sa.nb_null) #AIC is 434

sa.zi_null <- glmmTMB(Total_Tally~Treat_Type + (1|Site/Plot_No),
                       ziformula=~1,
                       family = nbinom2,
                       data=sapling_PIRI)
summary(sa.zi_null) #AIC is higher - need to learn more about how to fit/check the fit of models


# using DHARMa to check the model fit -------------------

sa_sim.resid <- simulateResiduals(sa.nb_null, n = 1000, plot = TRUE) #model differs significatly

testDispersion(sa_sim.resid)
testOutliers(sa_sim.resid)

plotResiduals(sa_sim.resid, sapling_PIRI$Treat_Type, xlab = "Treatment_Type", main=NULL)

plotResiduals(sa_sim.resid, sapling_PIRI$Site, xlab = "Site", main=NULL)

outliers(sa_sim.resid) #integer(0) is the given response???

testZeroInflation(sa_sim.resid) #this data appears to be zero inflated, so I have to figure out more about writing zero inflated code and then recheck

sa.zi_region <- glmmTMB(Total_Tally~Treat_Type + (1|Site/Plot_No),
                      zi=~Region,
                      family = nbinom2,
                      data=sapling_PIRI)
summary(sa.zi_region) #AIC is 467.7, more than null model, but null model is a poor fit
#nbinom1 as family has model convergence problems

sa.zi_sim.resid <- simulateResiduals(sa.zi_region, n=1000, plot = TRUE)

testZeroInflation(sa.zi_sim.resid) #holy shit, this looks good


# plot residuals against predictors/random variables/zi assignment
plotResiduals(sa.zi_sim.resid, sapling_PIRI$Treat_Type, xlab = "Treat_Type", main = NULL)

plotResiduals(sa.zi_sim.resid, sapling_PIRI$Site, xlab = "Site", main = NULL)

plotResiduals(sa.zi_sim.resid, sapling_PIRI$Region, xlab = "Region", main = NULL)

# post-hoc pairwise comparison -------------------
emmeans(sa.zi_region, pairwise~Treat_Type, adjust = "Tukey", type = "response") #not significant
```

























